Holomorphic field theories and higher algebra

### 2022-08-21 draft.tex
added more reminders to myself

\dedicatory{To Yuri Manin, with great admiration.}

@@

It is a remarkable fact, suggested by physicists, that there is also a map
\[
Y: {\rm Vac}[V] \to \End({\rm Vac}[V])[[z, z^{-1}]],
\]
known as the {\em vertex operator} or {\em state-field correspondence},
which is closely related to the $W[V]$-action.
The intuition behind $Y$ is that for any map $\phi: \CC^\times \to V$ and for any point $w \in \CC^\times$, 
there is a restriction of $\phi$ to a little disk around $w$.
That is, there is a restriction map $r_w: V[z,z^{-1}] \to V[[t]]$, where $t$ is the local coordinate $t = z-w$.
Hence there is a $w$-dependent map $\cO(V[[t]]) \to \cO(V[z,z^{-1}])$ by pulling back a function along $r_w$;
this map determines a $w$-dependent action of $\cO(V[[t]])$ on $\cO(V[[z]])$.
One can, in essence, quantize this map, by trying to extend the formulas of the $W[V]$-action,
and this leads to the map~$Y$.
If one axiomatizes the behavior of $Y$, one is led to the notion of a vertex algebra.

@@

Suppose now that $G$ is a complex Lie group and it acts on the symplectic complex vector space $V$ by holomorphic symplectomorphisms.
Then $G$ also acts on the algebraic loop space $V[z,z^{-1}]$ and also on formal loops $V((z))$ as a ``global'' symmetry:
the action is independent of $z$.
On the other hand, if we consider the infinitesimal action $\rho: \fg \to {\rm SympVect}(V)$,
there is, in fact, a natural extension to a ``local'' symmetry $\rho^{\rm loc}: \fg[z,z^{-1}] \to {\rm SympVect}(V((z)))$.
Geometrically, a $\fg$-valued function $x(z)$ acts pointwise (with respect to the formal disk) on the formal loop space.
(Although the algebraic loop space is not symplectic in a strict sense,
there is still a local symmetry as $\fg[z,z^{-1}]$ maps to vector fields on the algebraic loop space.)
This action factors through a map $H_\rho^{\rm loc}: \fg[z,z^{-1}] \to \cO(V((z)))$,
and hence is Hamiltonian.
Thus the loop algebra $\fg[z,z^{-1}]$ arises naturally as symmetries of a classical holomorphic field theory.

@@

We can use this algebra and its nice properties to define a kind of ``sphere algebra'' $\Map(\pAA^d,\fg)$ of algebraic maps from the punctured affine space (or the punctured formal disk) into $\fg$,
just as $\fg[z,z^{-1}]$ provides a kind of loop algebra.

@@

Associative algebras arise in quantum mechanics, we have argued, because the spacetime of a mechanical system is a line (``time'').
If we organize the observables (or operators) by when they happen in time,
it corresponds to writing them left-to-right as products in an algebra.
A natural question to ask then is what kind of higher algebraic structure captures how observables behave for theories in higher dimensions?
The shortest answer is ``factorization algebras,'' and we will describe and explore this notion in Section~\ref{sec fact alg},
but there are nice answers for two special classes of theories that we will develop now:
\begin{itemize}
\item For a topological field theory on $\RR^d$ (which we'll define in a moment),
the observables form an algebra over the {\em little $d$-disks operad}~$\Disc_d$ (sometimes also known as the $E_d$ operad). 
Equivalently, one says the observables are an $E_d$-algebra. 
This notion was introduced by algebraic topologists in the 1960s, 
so there is a rich body of work to draw upon.
\item For a holomorphic field theory on $\CC^d$,
the observables form an algebra over the $d$-polydiscs operad~$\PDisc_d$.
This notion was introduced only recently in~\cite{CG1}.
\end{itemize}
We will make the assertions above precise (in part by adding some natural but necessary hypotheses on a theory)
and develop them in parallel.
As a particularly illuminating example, we will explore the $d=1$ case rather closely and explain in what sense a vertex algebra is a kind of ``holomorphic associative algebra.''
We will also discuss the polydiscs algebras for which the higher Kac-Moody algebras are a Lie-algebraic shadow.

@@

The heuristic idea is straightforward.
Given a field theory $\cT$ on $\RR^d$, there is a vector space (or possibly dg vector space) $\Obs_\cT(D_r(x))$ of measurements one can make in the open disk $D_r(x)$ of radius $r$ around the point $x$.
Any measurement in a small disk should provide a measurement on a bigger disk,
so there is a map $\Obs_\cT(D_r(x)) \to \Obs_\cT(D_R(y))$ whenever $D_r(x) \subset D_R(y)$.
Thus we have a collection of vector spaces labeled by the disks in $\RR^d$,
and we have identified a class of unary operations determined by inclusion.

There should also be natural $n$-ary operations encoding how to combine observables from several distinct disks to give an observable on a big disk.
Start with $n$ disks $D_{r_1}(x_1)$, \dots , $D_{r_n}(x_n)$ whose closures are pairwise disjoint, and let $D_R(y)$ be a big disk containing all the smaller disks.
Then we want a map
\[
\Obs_\cT(D_{r_1}(x_1)) \otimes \cdots \otimes \Obs_\cT(D_{r_n}(x_n)) \to \Obs_\cT(D_R(y)).
\]
In the limit when the big disk is all of $\RR^d$ (i.e., $R = \infty$) and the input radii $r_i$ are very small, 
this map should encode the $n$-point functions beloved of physicists.

So far we have described a structure that observables should have for any field theory on~$\RR^d$.
For a topological field theory, we expect that the size of a disk does not matter:
for any inclusion $D_r(x) \subset D_R(y)$, the map $\Obs_\cT(D_r(x)) \to \Obs_\cT(D_R(y))$ is an equivalence.
This condition implies that we can always shrink the input radius as small as we want and we can always take the output radius to be a fixed value, say $R =1$.
In other words, the unary operations are parametrized by a point in the unit disk;
we have a local system of isomorphisms.

For $n$-ary operations we can likewise shrink the input radii as small as we want.
Thus the $n$-ary operations should be parametrized by the configuration space of $n$ points inside the unit disk.
(Rather, we can recover all $n$-ary operations knowing this information.)
We expect that these $n$-ary operations should also define a local system.

Our list of expectations is motivated by the examples of topological field theories that physicists typically give, rather than the Atiyah-Segal approach in terms of functors out of bordism categories.
Loosely speaking, a classical field theory (defined by a Lagrangian density, also known as an action functional) is ``topological'' if solutions to the equations of motion do not depend on the geometry of the spacetime manifold but only on its smooth structure (or other differential-topological features, like orientations).
A model example is Chern-Simons theory on oriented 3-manifolds for compact Lie group~$G$:
the space of solutions to the equations of motion on $M$ is the moduli space of flat $G$-bundles on $M$.
The observables for a classical field theory on $M$ means the commutative algebra of functions on solutions to the equations of motion on $M$.
In particular, for $M$ a disk, the size of the disk will not affect the algebra of observables,
and inclusions of disks leads to isomorphisms of observables.
Several of these features extend to the quantum field theories that arise by quantizing such topological classical theories;
the observables continue to form a vector space but are not commutative algebras anymore.
(See \cite{CG1, CG2} for a precise mathematical formulation of this idea, which is a natural extension of deformation quantization to higher dimensions.)

@@

\begin{dfn}
The {\em little $d$-disks operad} $\Disc_d$ is the topological operad whose space of $n$-ary operations is
\begin{align*}
\Disc_d(n) = \{ (x_1, \ldots, x_n; r_1, &\ldots, r_n)  \in D_1(0)^n \times (0,1)^n \;|\; \\
&\text{the closure of the disks $\overline{D_{r_i}(x_i)}$ is pairwise disjoint}\}.
\end{align*}
The composition $\mu \circ_i \nu$ is given by rescaling the unit disk that labels the output of $\nu$ to $r_i$ and then embedding it into the disk~$D_{r_i}(x_i)$.
\end{dfn}

\owen{Draw some 2d pictures and unpack a little. Mention pair of pants \& OPE} 
In this sense, it captures the dependence on the disks.

@@

\begin{dfn}
An {\em algebra over the little $d$-disks operad} (or little $d$-disks algebra or $E_d$-algebra) is a cochain complex $A$ that is an algebra over the dg operad~$C_*(\Disc_d)$.
\end{dfn}

We can state now encapsulate our expectations from above:
\begin{quote}
the observables of a topological field theory on $\RR^d$ are a little $d$-disks algebra.
\end{quote}
There is a setting where a precise version of this statement was proven, by Elliott and Safronov, for a mathematically well-defined class of Lagrangian field theories.

@@

\begin{thm}[\cite{EllSaf},\owen{precise location}]
Let $\cT$ be a classical field theory on $\RR^d$ in the sense of~\cite{CosBook},
and let its observables be constructed following~\cite{CG2}.
If the theory satisfies 
\begin{enumerate}
\item[(i)] $\cT$ is equivariant under the action of the translation group,
\item[(ii)] the action of the translation Lie algebra is homotopically trivialized, and
\item[(iii)] the extension of observables along an inclusion of disks is a quasi-isomorphism,
\end{enumerate}
then the observables form a little $d$-disks algebra.
\end{thm}

@@

This theorem applies to many well-known examples from physics, such as 
\begin{itemize}
\item the topological $B$-model of maps from an oriented surface into a Calabi-Yau manifold (constructed by Li and Li \cite{LiLi} in a setting where \cite{EllSaf} applies);
\item Chern-Simons gauge theory of flat $G$-bundles on an oriented 3-manifold (constructed, essentially, by Axelrod-Singer \cite{AxeSin} and Kontsevich \cite{KonECM} but articulated in this setting by Costello \cite{CosBook});
\item Rozansky-Witten theory of maps from an oriented 3-manifold into a complex manifold \cite{ChanLeungLi}; and
\item the Kapustin-Witten gauge theory for oriented 4-manifolds \cite{EGW},
\end{itemize}
among other examples of topological AKSZ theories, such as BF theories.

For physicists, we note that the proof uses a version of Witten descent.
Indeed, descent can be used to produce interesting operations in these disk algebras,
so that many familiar constructions from the physical approach to TFTs fit cleanly and easily into this framework.

@@

In the setting of a holomorphic field theory, we expect the observables to have a nontrivial dependence on the radius of a disk.
Consider, for example, that 
when $D_r(x) \subset D_R(y) \subset \CC$ are open disks in the plane,
the restriction map of holomorphic functions
\[
\cO(D_R(y))\to \cO(D_r(x))
\]
is not an isomorphism,
and the sheaf $\cO$ provides a fundamental model for how the solutions of the equations of motion should behave for a holomorphic field theory.
We will assume, however, that a holomorphic theory is equivariant under translations of $\CC^d$ and that the observables in a disk $D_r(X)$ are equivalent to the observables in a translate~$D_r(x+z)$.
Thus, we will have a colored operad whose colors are labeled by radii~$r \in (0,\infty)$,
as we merely have to specify the observables on a disk~$D_r(0)$.

We now introduce a small tweak.
Instead of working with disks, we work with polydisks: for $w = (w_1, \ldots, w_d) \in \CC^d$, the polydisk of radius $r$ around $w$~is
\[
PD_r(w) = \{ z \in \CC^d \;|\; |z_i - w_i| < r \text{ for all $i$}\}.
\]
We work with polydisks because it is easier to describe holomorphic functions on a polydisk than on a disk;
one can borrow results from single-variable complex analysis.
With this notion in hand, we introduce our main character.

\begin{dfn}
The {$d$-polydisks operad} $\PDisc_d$ is the colored operad in the category of complex manifolds where each positive real number $r \in (0,\infty)$ is a color 
and where there is a complex manifold of $n$-ary operations
\begin{align*}
\PDisc_d(r_1, \ldots, r_n| R) = \{ (w_1, \ldots,& w_n) \in PD_R(0)^n \;|\; 
\text{the closure of the polydisks $\overline{PD_{r_i}(w_i)}$}\\ 
&\text{ is pairwise disjoint and each is contained in $PD_R(0)$}\}
\end{align*}
for any list of radii $r_1,\ldots, r_n, R$.
This manifold is always an open subset of~$PD_R(0)^n$. 
(Note that the manifold is empty for many values.)
There is a composition
\[
\circ_i: \PDisc_d(r_1, \ldots, r_n| R) \times \PDisc_d(s_1, \ldots, s_m| S) \to \PDisc_d(r_1, \ldots, s_1, \ldots, s_m, r_{i+1}, \ldots, r_n| R)
\]
only when $r_i = S$, 
and in that case, the composition $\mu \circ_i \nu$ is given by embedding an output polydisk into an input polydisk of the same radius.
(Composition is not defined if the radii do not match.)
\end{dfn}

We are again thinking about configuration spaces --- here of polydisks inside larger polydisks --- but we are remembering the complex structure explicitly.
To describe observables, we want to produce a dg colored operad that encodes the holomorphic nature of this situation.
We do {\em not} want to take singular chains, which would only remember the homotopy type;
instead, we choose to take the Dolbeault complex of each configuration space.
Note, however, that taking the Dolbeault complex is a contravariant functor,
so we get a {\em co}\/operad rather than an operad.
This notion bears the same relationship to operad as a coalgebra does to an algebra;
one simply reverses all the arrows and asks for an appropriate variant of the relations.
For instance, there is a composition
\begin{align}
\label{pdcirc}
\circ_i: \Omega^{0,*}(\PDisc_d(r_1, \ldots, s_1, &\ldots, s_m, r_{i+1}, \ldots, r_n| R)) \to \\
&\Omega^{0,*}(\PDisc_d(r_1, \ldots, r_n| R)) \otimes \Omega^{0,*}(\PDisc_d(s_1, \ldots, s_m| S))
\end{align}
\owen{I don't know how to eliminate the two numbers}
when $r_i = S$, and in that case, the composition $\mu \circ_i \nu$ is given by embedding an output polydisk into an input polydisk of the same radius.
(Composition is not defined if the radii do not match.)

@@

\begin{dfn}
The {\em $d$-polydisks dg cooperad} is the colored operad in the category of dg vector spaces where each positive real number $r \in (0,\infty)$ is a color 
and where there is a dg vector space of $n$-ary cooperations
\[
\Omega^{0,*}(\PDisc_d(r_1, \ldots, r_n| R)) 
\]
for any list of radii $r_1,\ldots, r_n, R$.
There is a composition
\[
\circ_i: \Omega^{0,*}(\PDisc_d(r_1, \ldots, s_1, \ldots, s_m, r_{i+1}, \ldots, r_n| R))\to \Omega^{0,*}(\PDisc_d(r_1, \ldots, r_n| R)) \otimes \Omega^{0,*}(\PDisc_d(s_1, \ldots, s_m| S))
\]
only when $r_i = S$, 
and in that case, the composition $\mu \circ_i \nu$ is given by embedding an output polydisk into an input polydisk of the same radius.
(Composition is not defined if the radii do not match.)
\end{dfn}

@@

In the case $d=1$, it is possible to relate polydisks algebras to vertex algebras.
More precisely, when the polydisk algebra has a nice action of the rotation group~$U(1)$,
one can extract a vertex algebra: 
see Chapter 5 of~\cite{CG1} for a detailed treatment and examples like the free $\beta\gamma$ system and the Kac-Moody vertex algebras.
Here we will analyze the cohomology of the cooperad and what kind of information it encodes, with a view towards motivating the relationship to vertex algebras.

@@

It is easy to see
\[
{\rm Conf}_n(\CC) \cong \CC \times (\CC^\times)^{n-1}
\]
by sending a configuration $(w_1,\ldots, w_n)$ to $(w_1, w_2 - w_1, \ldots, w_n-w_{n-1})$.
(We could replace $w_1$ by any $w_j$, or even the sum of them all.)
Hence, the Dolbeault cohomology consists of holomorphic functions on this space;
there is no higher cohomology.
Inside the holomorphic functions, as a dense but understandable subalgebra, are the algebraic functions:
\[
\CC[w_1, (w_1 - w_2)^{\pm 1}, \ldots, (w_n - w_{n-1})^{\pm 1}] \subset H^{0,*}({\rm Conf}_n(\CC)).
\]
Here $(w_j - w_{j-1})^{\pm 1}$ means that the function $1/(w_j - w_{j-1})$ is also a generator.
Each such algebraic function provides an operation on $H^*(A)$, where $A$ denotes a polydisk algebra.

@@

\owen{I think I should swap the location of $a$ and $b$ to match VOA conventions.}

\subsubsection{Polydisks algebras from holomorphic field theories}

@@

\begin{thm}[\cite{CG1},\owen{precise location}]
Let $\cT$ be a classical field theory on $\CC^d$ in the sense of~\cite{CosBook},
and let its observables be constructed following~\cite{CG2}.
If the theory satisfies 
\begin{enumerate}
\item[(i)] $\cT$ is equivariant under the action of the translation group and
\item[(ii)] the action of the Lie algebra $\ft_{\zbar}$ is homotopically trivialized,
\end{enumerate}
then the observables form a $d$-polydisks algebra.
\end{thm}

@@

\begin{eg}\label{WeissAsExponential}
For a smooth $n$-manifold $M$, there are several simple ways to construct a Weiss cover for $M$. One construction is simply to take the collection of open sets in $M$ diffeomorphic to a disjoint union of finitely many copies of the open $n$-disk. Alternatively, the collection of opens $\{ M \setminus q \mcol q \in M\}$ is a Weiss cover of $M$. One can also produce a Weiss cover of a manifold with countably many elements. Fix a Riemannian metric on $M$. Pick a collection of points $\{q_n\in M \mcol n \in \NN\}$ such that the union of the unit ``discs'' $D_1(q_n) = \{ q \in M \mcol \d(q_n,q) < 1\}$ covers $M$. (Such an open may not be homeomorphic to a disc if the injectivity radius of $q_n$ is less than 1.)  Consider the collection of all ``discs'' $\fD = \{ D_{1/m}(q_n) \mcol m \in \NN, n \in \NN\}$.  The collection of all finite, pairwise disjoint union of elements of $\fD$ is a countable Weiss cover.~\hfill $\Diamond$ 
\end{eg}

@@

\subsection{Algebras over the little disks operad}

>>>>

\dedicatory{To Yuri Manin, with gratitude and admiration.}

@@

It is a remarkable fact, suggested by physicists, that there is also a map
\[
Y: {\rm Vac}[V] \to \End({\rm Vac}[V])[[z, z^{-1}]],
\]
known as the {\em vertex operator} or {\em state-field correspondence},
which is closely related to the $W[V]$-action.
The intuition behind $Y$ is that for any map $\phi: \CC^\times \to V$ and for any point $w \in \CC^\times$, 
there is a restriction of $\phi$ to a little disc around $w$.
That is, there is a restriction map $r_w: V[z,z^{-1}] \to V[[t]]$, where $t$ is the local coordinate $t = z-w$.
Hence there is a $w$-dependent map $\cO(V[[t]]) \to \cO(V[z,z^{-1}])$ by pulling back a function along $r_w$;
this map determines a $w$-dependent action of $\cO(V[[t]])$ on $\cO(V[[z]])$.
One can, in essence, quantize this map, by trying to extend the formulas of the $W[V]$-action,
and this leads to the map~$Y$.
If one axiomatizes the behavior of $Y$, one is led to the notion of a vertex algebra.

@@

Suppose now that $G$ is a complex Lie group and it acts on the symplectic complex vector space $V$ by holomorphic symplectomorphisms.
Then $G$ also acts on the algebraic loop space $V[z,z^{-1}]$ and also on formal loops $V((z))$ as a ``global'' symmetry:
the action is independent of $z$.
On the other hand, if we consider the infinitesimal action $\rho: \fg \to {\rm SympVect}(V)$,
there is, in fact, a natural extension to a ``local'' symmetry $\rho^{\rm loc}: \fg[z,z^{-1}] \to {\rm SympVect}(V((z)))$.
Geometrically, a $\fg$-valued function $x(z)$ acts pointwise (with respect to the formal disc) on the formal loop space.
(Although the algebraic loop space is not symplectic in a strict sense,
there is still a local symmetry as $\fg[z,z^{-1}]$ maps to vector fields on the algebraic loop space.)
This action factors through a map $H_\rho^{\rm loc}: \fg[z,z^{-1}] \to \cO(V((z)))$,
and hence is Hamiltonian.
Thus the loop algebra $\fg[z,z^{-1}]$ arises naturally as symmetries of a classical holomorphic field theory.

@@

We can use this algebra and its nice properties to define a kind of ``sphere algebra'' $\Map(\pAA^d,\fg)$ of algebraic maps from the punctured affine space (or the punctured formal disc) into $\fg$,
just as $\fg[z,z^{-1}]$ provides a kind of loop algebra.

@@

Associative algebras arise in quantum mechanics, we have argued, because the spacetime of a mechanical system is a line (``time'').
If we organize the observables (or operators) by when they happen in time,
it corresponds to writing them left-to-right as products in an algebra.
A natural question to ask then is what kind of higher algebraic structure captures how observables behave for theories in higher dimensions?
The shortest answer is ``factorization algebras,'' and we will describe and explore this notion in Section~\ref{sec fact alg},
but there are nice answers for two special classes of theories that we will develop now:
\begin{itemize}
\item For a topological field theory on $\RR^d$ (which we'll define in a moment),
the observables form an algebra over the {\em little $d$-discs operad}~$\Disc_d$ (sometimes also known as the $E_d$ operad). 
Equivalently, one says the observables are an $E_d$-algebra. 
This notion was introduced by algebraic topologists in the 1960s, 
so there is a rich body of work to draw upon.
\item For a holomorphic field theory on $\CC^d$,
the observables form an algebra over the $d$-polydiscs operad~$\PDisc_d$.
This notion was introduced only recently in~\cite{CG1}.
\end{itemize}
We will make the assertions above precise (in part by adding some natural but necessary hypotheses on a theory)
and develop them in parallel.
As a particularly illuminating example, we will explore the $d=1$ case rather closely and explain in what sense a vertex algebra is a kind of ``holomorphic associative algebra.''
We will also discuss the polydiscs algebras for which the higher Kac-Moody algebras are a Lie-algebraic shadow.

@@

The heuristic idea is straightforward.
Given a field theory $\cT$ on $\RR^d$, there is a vector space (or possibly dg vector space) $\Obs_\cT(D_r(x))$ of measurements one can make in the open disc $D_r(x)$ of radius $r$ around the point $x$.
Any measurement in a small disc should provide a measurement on a bigger disc,
so there is a map $\Obs_\cT(D_r(x)) \to \Obs_\cT(D_R(y))$ whenever $D_r(x) \subset D_R(y)$.
Thus we have a collection of vector spaces labeled by the discs in $\RR^d$,
and we have identified a class of unary operations determined by inclusion.

There should also be natural $n$-ary operations encoding how to combine observables from several distinct discs to give an observable on a big disc.
Start with $n$ discs $D_{r_1}(x_1)$, \dots , $D_{r_n}(x_n)$ whose closures are pairwise disjoint, and let $D_R(y)$ be a big disc containing all the smaller discs.
Then we want a map
\[
\Obs_\cT(D_{r_1}(x_1)) \otimes \cdots \otimes \Obs_\cT(D_{r_n}(x_n)) \to \Obs_\cT(D_R(y)).
\]
In the limit when the big disc is all of $\RR^d$ (i.e., $R = \infty$) and the input radii $r_i$ are very small, 
this map should encode the $n$-point functions beloved of physicists.

So far we have described a structure that observables should have for any field theory on~$\RR^d$.
For a topological field theory, we expect that the size of a disc does not matter:
for any inclusion $D_r(x) \subset D_R(y)$, the map $\Obs_\cT(D_r(x)) \to \Obs_\cT(D_R(y))$ is an equivalence.
This condition implies that we can always shrink the input radius as small as we want and we can always take the output radius to be a fixed value, say $R =1$.
In other words, the unary operations are parametrized by a point in the unit disc;
we have a local system of isomorphisms.

For $n$-ary operations we can likewise shrink the input radii as small as we want.
Thus the $n$-ary operations should be parametrized by the configuration space of $n$ points inside the unit disc.
(Rather, we can recover all $n$-ary operations knowing this information.)
We expect that these $n$-ary operations should also define a local system.

Our list of expectations is motivated by the examples of topological field theories that physicists typically give, rather than the Atiyah-Segal approach in terms of functors out of bordism categories.
Loosely speaking, a classical field theory (defined by a Lagrangian density, also known as an action functional) is ``topological'' if solutions to the equations of motion do not depend on the geometry of the spacetime manifold but only on its smooth structure (or other differential-topological features, like orientations).
A model example is Chern-Simons theory on oriented 3-manifolds for compact Lie group~$G$:
the space of solutions to the equations of motion on $M$ is the moduli space of flat $G$-bundles on $M$.
The observables for a classical field theory on $M$ means the commutative algebra of functions on solutions to the equations of motion on $M$.
In particular, for $M$ a disc, the size of the disc will not affect the algebra of observables,
and inclusions of discs leads to isomorphisms of observables.
Several of these features extend to the quantum field theories that arise by quantizing such topological classical theories;
the observables continue to form a vector space but are not commutative algebras anymore.
(See \cite{CG1, CG2} for a precise mathematical formulation of this idea, which is a natural extension of deformation quantization to higher dimensions.)

@@

\begin{dfn}
The {\em little $d$-discs operad} $\Disc_d$ is the topological operad whose space of $n$-ary operations is
\begin{align*}
\Disc_d(n) = \{ (x_1, \ldots, x_n; r_1, &\ldots, r_n)  \in D_1(0)^n \times (0,1)^n \;|\; \\
&\text{the closure of the discs $\overline{D_{r_i}(x_i)}$ is pairwise disjoint}\}.
\end{align*}
The composition $\mu \circ_i \nu$ is given by rescaling the unit disc that labels the output of $\nu$ to $r_i$ and then embedding it into the disc~$D_{r_i}(x_i)$.
\end{dfn}

\owen{Draw some 2d pictures and unpack a little. Mention pair of pants \& OPE} 
In this sense, it captures the dependence on the discs.

@@

\begin{dfn}
An {\em algebra over the little $d$-discs operad} (or little $d$-discs algebra or $E_d$-algebra) is a cochain complex $A$ that is an algebra over the dg operad~$C_*(\Disc_d)$.
\end{dfn}

\owen{Perhaps mention classic examples like $n$-fold loop spaces or Hochschild cohomology?}

We can state now encapsulate our expectations from above:
\begin{quote}
the observables of a topological field theory on $\RR^d$ are a little $d$-discs algebra.
\end{quote}
There is a setting where a precise version of this statement was proven, by Elliott and Safronov, for a mathematically well-defined class of Lagrangian field theories.

@@

\begin{thm}[\cite{EllSaf},\owen{precise location}]
Let $\cT$ be a classical field theory on $\RR^d$ in the sense of~\cite{CosBook},
and let its observables be constructed following~\cite{CG2}.
If the theory satisfies 
\begin{enumerate}
\item[(i)] $\cT$ is equivariant under the action of the translation group,
\item[(ii)] the action of the translation Lie algebra is homotopically trivialized, and
\item[(iii)] the extension of observables along an inclusion of discs is a quasi-isomorphism,
\end{enumerate}
then the observables form a little $d$-discs algebra.
\end{thm}

@@

This theorem applies to many well-known examples from physics, such as 
\begin{itemize}
\item the topological $B$-model of maps from an oriented surface into a Calabi-Yau manifold (constructed by Li and Li \cite{LiLi} in a setting where \cite{EllSaf} applies);
\item Chern-Simons gauge theory of flat $G$-bundles on an oriented 3-manifold (constructed, essentially, by Axelrod-Singer \cite{AxeSin} and Kontsevich \cite{KonECM} but articulated in this setting by Costello \cite{CosBook});
\item Rozansky-Witten theory of maps from an oriented 3-manifold into a complex manifold \cite{ChanLeungLi}; and
\item the Kapustin-Witten gauge theory for oriented 4-manifolds \cite{EGW},
\end{itemize}
among other examples of topological AKSZ theories, such as BF theories.
In all these examples, the underlying elliptic complex is a version of the de Rham complex.

For physicists, we note that the proof uses a version of Witten descent.
Indeed, descent can be used to produce interesting operations in these disc algebras,
so that many familiar constructions from the physical approach to TFTs fit cleanly and easily into this framework.

@@

In the setting of a holomorphic field theory, we expect the observables to have a nontrivial dependence on the radius of a disc.
Consider, for example, that 
when $D_r(x) \subset D_R(y) \subset \CC$ are open discs in the plane,
the restriction map of holomorphic functions
\[
\cO(D_R(y))\to \cO(D_r(x))
\]
is not an isomorphism,
and the sheaf $\cO$ provides a fundamental model for how the solutions of the equations of motion should behave for a holomorphic field theory.
We will assume, however, that a holomorphic theory is equivariant under translations of $\CC^d$ and that the observables in a disc $D_r(X)$ are equivalent to the observables in a translate~$D_r(x+z)$.
Thus, we will have a colored operad whose colors are labeled by radii~$r \in (0,\infty)$,
as we merely have to specify the observables on a disc~$D_r(0)$.

We now introduce a small tweak.
Instead of working with discs, we work with polydiscs: for $w = (w_1, \ldots, w_d) \in \CC^d$, the polydisc of radius $r$ around $w$~is
\[
PD_r(w) = \{ z \in \CC^d \;|\; |z_i - w_i| < r \text{ for all $i$}\}.
\]
We work with polydiscs because it is easier to describe holomorphic functions on a polydisc than on a disc;
one can borrow results from single-variable complex analysis.
With this notion in hand, we introduce our main character.

\begin{dfn}
The {$d$-polydiscs operad} $\PDisc_d$ is the colored operad in the category of complex manifolds where each positive real number $r \in (0,\infty)$ is a color 
and where there is a complex manifold of $n$-ary operations
\begin{align*}
\PDisc_d(r_1, \ldots, r_n| R) = \{ (w_1, \ldots,& w_n) \in PD_R(0)^n \;|\; 
\text{the closure of the polydiscs $\overline{PD_{r_i}(w_i)}$}\\ 
&\text{ is pairwise disjoint and each is contained in $PD_R(0)$}\}
\end{align*}
for any list of radii $r_1,\ldots, r_n, R$.
This manifold is always an open subset of~$PD_R(0)^n$. 
(Note that the manifold is empty for many values.)
There is a composition
\[
\circ_i: \PDisc_d(r_1, \ldots, r_n| R) \times \PDisc_d(s_1, \ldots, s_m| S) \to \PDisc_d(r_1, \ldots, s_1, \ldots, s_m, r_{i+1}, \ldots, r_n| R)
\]
only when $r_i = S$, 
and in that case, the composition $\mu \circ_i \nu$ is given by embedding an output polydisc into an input polydisc of the same radius.
(Composition is not defined if the radii do not match.)
\end{dfn}

We are again thinking about configuration spaces --- here of polydiscs inside larger polydiscs --- but we are remembering the complex structure explicitly.
To describe observables, we want to produce a dg colored operad that encodes the holomorphic nature of this situation.
We do {\em not} want to take singular chains, which would only remember the homotopy type;
instead, we choose to take the Dolbeault complex of each configuration space.
Note, however, that taking the Dolbeault complex is a contravariant functor,
so we get a {\em co}\/operad rather than an operad.
This notion bears the same relationship to operad as a coalgebra does to an algebra;
one simply reverses all the arrows and asks for an appropriate variant of the relations.
For instance, there is a composition
\begin{align}
\label{pdcirc}
\circ_i: \Omega^{0,*}(\PDisc_d(r_1, \ldots, s_1, &\ldots, s_m, r_{i+1}, \ldots, r_n| R)) \to \\
&\Omega^{0,*}(\PDisc_d(r_1, \ldots, r_n| R)) \otimes \Omega^{0,*}(\PDisc_d(s_1, \ldots, s_m| S))
\end{align}
\owen{I don't know how to eliminate the two numbers}
when $r_i = S$, and in that case, the composition $\mu \circ_i \nu$ is given by embedding an output polydisc into an input polydisc of the same radius.
(Composition is not defined if the radii do not match.)

@@

\begin{dfn}
The {\em $d$-polydiscs dg cooperad} is the colored operad in the category of dg vector spaces where each positive real number $r \in (0,\infty)$ is a color 
and where there is a dg vector space of $n$-ary cooperations
\[
\Omega^{0,*}(\PDisc_d(r_1, \ldots, r_n| R)) 
\]
for any list of radii $r_1,\ldots, r_n, R$.
There is a composition
\[
\circ_i: \Omega^{0,*}(\PDisc_d(r_1, \ldots, s_1, \ldots, s_m, r_{i+1}, \ldots, r_n| R))\to \Omega^{0,*}(\PDisc_d(r_1, \ldots, r_n| R)) \otimes \Omega^{0,*}(\PDisc_d(s_1, \ldots, s_m| S))
\]
only when $r_i = S$, 
and in that case, the composition $\mu \circ_i \nu$ is given by embedding an output polydisc into an input polydisc of the same radius.
(Composition is not defined if the radii do not match.)
\end{dfn}

@@

In the case $d=1$, it is possible to relate polydiscs algebras to vertex algebras.
More precisely, when the polydisc algebra has a nice action of the rotation group~$U(1)$,
one can extract a vertex algebra: 
see Chapter 5 of~\cite{CG1} for a detailed treatment and examples like the free $\beta\gamma$ system and the Kac-Moody vertex algebras.
Here we will analyze the cohomology of the cooperad and what kind of information it encodes, with a view towards motivating the relationship to vertex algebras.

@@

It is easy to see
\[
{\rm Conf}_n(\CC) \cong \CC \times (\CC^\times)^{n-1}
\]
by sending a configuration $(w_1,\ldots, w_n)$ to $(w_1, w_2 - w_1, \ldots, w_n-w_{n-1})$.
(We could replace $w_1$ by any $w_j$, or even the sum of them all.)
Hence, the Dolbeault cohomology consists of holomorphic functions on this space;
there is no higher cohomology.
Inside the holomorphic functions 
\[
\cO({\rm Conf}_n(\CC)) = H^{0,*}({\rm Conf}_n(\CC)),
\] 
as a dense but understandable subalgebra, are the algebraic functions:
\[
\cO_{\rm alg}({\rm Conf}_n(\CC)) = \CC[w_1, (w_1 - w_2)^{\pm 1}, \ldots, (w_n - w_{n-1})^{\pm 1}].
\]
Here $(w_j - w_{j-1})^{\pm 1}$ means that the function $1/(w_j - w_{j-1})$ is a generator, 
as is $(w_j - w_{j-1})$.
Each such algebraic function provides an operation on $H^*(A)$, where $A$ denotes a polydisc algebra.

@@

\owen{I think I should swap the location of $a$ and $b$ to match VOA conventions. I want to explain that taking the extremal radii means we cannot compose, which leads to the complicated VOA formulas.}

\subsubsection{Revisiting the higher Kac-Moody algebras}

\owen{I want to mention here that we have a polydisc version}

\subsubsection{Polydiscs algebras from holomorphic field theories}

@@

\begin{thm}[\cite{CG1},\owen{precise location}]
Let $\cT$ be a classical field theory on $\CC^d$ in the sense of~\cite{CosBook},
and let its observables be constructed following~\cite{CG2}.
If the theory satisfies 
\begin{enumerate}
\item[(i)] $\cT$ is equivariant under the action of the translation group and
\item[(ii)] the action of the Lie algebra $\ft_{\zbar}$ is homotopically trivialized,
\end{enumerate}
then the observables form a $d$-polydiscs algebra.
\end{thm}

@@

\begin{eg}\label{WeissAsExponential}
For a smooth $n$-manifold $M$, there are several simple ways to construct a Weiss cover for $M$. One construction is simply to take the collection of open sets in $M$ diffeomorphic to a disjoint union of finitely many copies of the open $n$-disc. Alternatively, the collection of opens $\{ M \setminus q \mcol q \in M\}$ is a Weiss cover of $M$. One can also produce a Weiss cover of a manifold with countably many elements. Fix a Riemannian metric on $M$. Pick a collection of points $\{q_n\in M \mcol n \in \NN\}$ such that the union of the unit ``discs'' $D_1(q_n) = \{ q \in M \mcol \d(q_n,q) < 1\}$ covers $M$. (Such an open may not be homeomorphic to a disc if the injectivity radius of $q_n$ is less than 1.)  Consider the collection of all ``discs'' $\fD = \{ D_{1/m}(q_n) \mcol m \in \NN, n \in \NN\}$.  The collection of all finite, pairwise disjoint union of elements of $\fD$ is a countable Weiss cover.~\hfill $\Diamond$ 
\end{eg}

@@

\subsection{Algebras over the little discs operad}
